---
title: "PGH Traffic Speed Dashboard Visualization"
author: "Truc Viet Le"
date: "01/04/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the high-level description of the notebook.

Let's see what files are there in the input directory.

```{r echo=FALSE, message=FALSE}
list.files('./speed_app/data/')
```

## Read the Input Data

First, the road segment data.

```{r echo=FALSE, message=FALSE}
df_segment = read.csv('./speed_app/data/pgh_segments.csv')
paste('Total number of segments =', nrow(df_segment))
```

Then the training and prediction data sets.

```{r echo=FALSE, message=FALSE}
get_input_data <- function(is_weekday, select_time, is_training=TRUE) {
  if(is_training) {
    input_data <- read.csv("./speed_app/data/pgh_train.csv")
  } else {
    input_data <- read.csv("./speed_app/data/pgh_pred.csv")
  }
  
  if(is_weekday) {
    input_data <- subset(input_data, is.weekday==1)
  } else {
    input_data <- subset(input_data, is.weekday==0)
  }
  
  time_values <- unique(input_data$time)
  if(select_time %in% time_values) {
    input_data <- subset(input_data, time==select_time)
  } else {
    stop('The select time is invalid: ', select_time)
  }
  
  return(input_data)
}

is_weekday <- sample(c(FALSE, TRUE), size = 1)
if(is_weekday) {
  paste('Speed data for:', 'Weekday')
} else {
  paste('Speed data for:', 'Weekend')
}

select_time <- sample(c(8, 14, 20), size = 1)
paste('Select time =', select_time)

df_train <- get_input_data(is_weekday, select_time, is_training = TRUE)
paste('Number of observed speed values =', nrow(df_train))

df_pred <- get_input_data(is_weekday, select_time, is_training = FALSE)
paste('Number of predicted speed values =', nrow(df_pred))
```

## Make the Visualization

Create a new visualization data frame by combining predicted speed values with the segment data.

```{r echo=FALSE, message=FALSE}
speed_data <- data.frame(from.x = df_segment$from.x,
                         from.y = df_segment$from.y,
                         to.x = df_segment$to.x,
                         to.y = df_segment$to.y,
                         speed = df_pred$speed)
print(dim(speed_data))
```

Create spatial lines from the speed data and assign color values to them

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(sp)
library(scales)
library(leaflet)

convertSPLines <- function(input.data=data.frame()) {
  ## Converts an input segment data frame into spatial lines data.
  
  ## Specify the coordinates of the ending points
  begin.coord <- data.frame(lon = input.data$from.x, lat = input.data$from.y)
  end.coord <- data.frame(lon = input.data$to.x, lat = input.data$to.y)
  
  ## Road segment attributes
  speed <- round(input.data$speed, 2)
  attribute.data <- data.frame(speed = speed, stringsAsFactors = FALSE)
  
  l <- vector("list", nrow(begin.coord))
  for (i in seq_along(l)) {
    l[[i]] <- Lines(list(Line(rbind(begin.coord[i, ], end.coord[i,]))), as.character(i))
  }
  output <- SpatialLines(l)
  final.output <- SpatialLinesDataFrame(output, data = attribute.data)
  
  return(final.output)
}

# Convert speed data into spatial lines for drawing
speed_segment <- convertSPLines(speed_data)
print('Conversion completed!')

# Create the speed color legend
pal <- colorNumeric("RdYlGn", domain = NULL, na.color = "#808080")
```

Put visualization on the map

```{r echo=FALSE, message=FALSE}
# Pre-defined constants: Central coordinates of PGH
PGH_LON <- -79.98606 # longitude coordinate
PGH_LAT <- 40.44796 # latitude coordinate

my_viz <- leaflet(data = speed_segment) %>%
  addTiles(urlTemplate = "//{s}.tiles.mapbox.com/v3/jcheng.map-5ebohr46/{z}/{x}/{y}.png",
           attribution = 'Maps by <a href="http://www.mapbox.com/">Mapbox</a>') %>%
  setView(lng = PGH_LON, lat = PGH_LAT, zoom = 14) %>%
  addPolylines(color = ~pal(speed),
               weight = 3, opacity = 0.90) %>%
  addLegend("bottomleft", pal=pal, values=~speed, title='Speed (mph)',
            opacity = 0.90)
my_viz
```
